Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-07-04 WRITEUP/ms.tex
into conclusions

The most massive structures ($M_{\rm gas}>$10$^7$\,\Msun) correspond
to the molecular disk of the galaxy (see e.g., \Fig{MCC}), which is
identified as a single component at low $n_{\rm cut}$. This main disk
component
% mm2 occupy the top right   [left not right!]
          occupies the top left
corner of \Fig{stellarRatio16}, with a low gas mass fraction of
$f_{\rm gas}< 5\%$. The velocity dispersion of such a structure is
dominated by bulk
% mm2 motions,
rotational motion,
as shown in \Fig{vv}, where we compare the velocity dispersions of MCCs resulting from bulk ($\sigma_{\rm bulk}$) versus non-thermal turbulent motions ($\sigma_{\rm NT}$) in the accreting and quiescent phases of \flower. The accreting phase of \flower displays a disturbed morphology, whereas the quiescent phase displays a disk-like morphology (\Fig{MCC}).
%
Excluding the structure corresponding to the molecular disk, the
velocity dispersions of most MCCs in the accreting phase of \flower
are dominated by turbulent motions, 
% mm2 . On the other hand, [the first hand is missing]
  while  in the quiescent disk-like phase, $\sigma_{\rm bulk}$ is at
  least comparable to $\sigma_{\rm NT}$ for almost half the MCCs.
%mm2 in the quiescent disk-like phase.
%
For these reasons, the most massive MCCs are excluded in the discussion of the dynamics (\Sec{diss}), as they are dominated by large-scale shear; however, we include them in the scaling relation plots (\Fig{larsons_single} and \Fig{alpha16-28}), as would be done for an observational study.

@@

While the virial parameter defined as Equation~\ref{eqn:alpha} is typically used in observational studies, as they are designed to probe structures that are composed of mainly molecular gas only, we calculate the virial parameter of MCCs including the influence from the stellar component (i.e., $\alpha_{\rm vir, tot}$ via Equation~\ref{eqn:alpha_tot}) given the high stellar-to-gas mass ratios of some molecular structures identified (see \Fig{stellarRatio16}).
% 
The biggest MCC identified at low $n_{\rm cut}$ and with the highest stellar-to-gas mass ratio ($\gtrsim$\,1; \Fig{stellarRatio16}) corresponds to the central main 
disk of \flower (see top panels of \Fig{MCC} and \Sec{distribution}). The large amount of stellar mass already assembled in \flower
explains the low {\em total} virial parameter seen in the middle left panel of \Fig{larsons_single}, where the stellar gravitational potential 
influences the overall stability. On the other hand, the MCCs in this phase are mostly stable, with 
virial parameter comparable to some of the molecular structures
observed in the Milky Way,
% mm2 but the former is
     though the MCCs are
     more massive.

@@

%mm2 [fiddled with tenses and plurals]
If constant column density (Larson's third relation) were truly a
fundamental property of molecular clouds,
% or the roughly linear relationship between $\rho$ and $R$ of MCCs.
% mm2 (i.e., 
   and if virial equilibrium were a universal property of molecular clouds with negligible surface pressure, 
one would expect a single point in the $\sigma_{\rm
  gas}^2/R$--$\Sigma$ relation at
% mm a given gas
    that
mass surface density $\Sigma$. 
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental property of 
molecular clouds
% mm2 [maybe this is excessive, but if we are discussing the issue it
% seems worth citing.]
     in agreement with \citet{Heyer09a}.
Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[see e.g.,][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

We investigate possible variations in the dynamics of the molecular structures of \flower 
by adopting different
% mm2 sets
    values
of $n_{\rm cut}$ % in identifying the MCCs
to test the robustness of our results against the choice of density threshold. 
That is, how sensitively are the structure properties, and thus, the 
results presented in \Sec{singless}, dependent on the choice of density thresholds?

@@

In this section, we compare the observable quantities of MCCs obtained from the simulations (e.g., $\sigma_{\rm gas}$, $R$, $M_{\rm gas}$, $\Sigma_{\rm gas}$) with those of molecular gas structures from \obs at lower redshift.
%%
%We exclude the biggest MCC identified in each evolutionary stage, which correspond to the main disk of \flower (i.e., the top right cluster of points in the linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}).
The velocity dispersion of MCCs in \flower is similar to those observed in $z$\ssim2 spatially resolved studies of gas-rich star-forming galaxies, 
spanning a range of $\sigma_{\rm gas}\simeq$\,20$-$80\,\kms (\Fig{larsons_single}; see e.g., \citealt{Swinbank11a}). 
However, the sizes of MCCs are almost a factor of two smaller. This
% mm2 likely suggest
    suggests 
that the sizes of \highz molecular gas structures reported in the literature
are
% mm2
   probably
limited by the spatial resolution. 
%
The velocity dispersions of some MCCs are also comparable to the most turbulent gas clouds observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a, Leroy15a}), which lie along the locus of $\sigma_{\rm gas}\propto R^{0.56}$. Such high velocity dispersion 
in \flower is dominated by non-thermal energy (\Fig{vv}) and results from the injection of kinetic energy from recent \SF as \flower is assembling its stellar mass (given the high Mach numbers of MCCs; \Sec{distribution}). In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2, \flower has assembled a stellar mass of $M_\star$\eq7.5\E{9}\,\Msun.
%mm2
   This contrasts with \citet{Krumholz18}, who show that numerical models and analytic arguments point to
   feedback generally not being able to produce velocity dispersions of more than 10--20 km~s$^{-1}$, with
   higher velocities likely being produced by Toomre instability driven radial inflow.

The pressure of MCCs identified at the highest $n_{\rm cut}$ ranges in $P/k_B\simeq10^{7-9}$\,{\rm K}\,{\rm cm}$^{-3}$, with a median of $10^{7.6}$\,\,{\rm K}\,{\rm cm}$^{-3}$. Such high pressure is comparable to those observed in local ULIRGs; however, the molecular clouds in the latter are concentrated within their central regions \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs of \flower are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{Kohandel19a} and higher velocity accretion/SN-driven outflows \citep{Gallerani18a}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2. For instance, \citet{Swinbank11a} report ISM pressure of $P/k_B\sim$10$^7$\,K\,\cc and molecular gas mass of $M_{\rm gas}\simeq10^{8-9}$\,\Msun in the star-forming regions of a gas-rich starburst galaxy at \z\eq2.3 based on spatially resolved CO line \obs.

% different evolutionary stages
We do not find any major quantitative differences in the MCCs of \flower between its accreting and starburst phase with respect to those observed in the nearby Universe (see \Fig{larsons_single}), which are separated by $\simeq$300\,Myr. However, MCCs in the \SB phase have lower $\alpha_{\rm vir}$ compared to the accreting phase. This is expected for \SF to proceed.

@@

Close resemblance of the $Q_{\rm eff}$ and $Q_{\star}$ maps indicates that contributions from the stellar component play an important role in governing the stability of the MCCs against perturbations. This can be understood since stars in \flower dominates the central part of the galaxy in mass, so their gravitational potential provides a non-negligible contribution to the instability. In fact, this is also illustrated by the differences seen between $\alpha_{\rm vir}$ and $\alpha_{\rm vir, tot}$ 
of the most massive structure (i.e., the gas disk of \flower; see \Fig{larsons_single}).
Similarly, contribution from the thickness of the disk is important in \flower since its disk is warped and has a scale height-to-radius ratio of $h/r_{\rm gal}$\ssim150\,pc/1\,kpc\,$\simeq$\,0.15.
% 
That is, some MCCs are found in regions of $Q_{\rm gas}$\ssim1, which is consistent with the expectation that they correspond to regions of high surface densities that are gravitationally unstable. Note, however, that when including the stabilizing effects due to the stellar potential of \flower and the thickness of its disk (via $\sigma_z$, i.e., $Q_{\rm eff}$; see Equation~\ref{eqn:q_eff}), some of these MCCs are consistent with $Q_{\rm eff} > 1$. 
%
This demonstrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures. This consideration is especially relevant for the relatively evolved and enriched systems at high redshift, that are preferentially being imaged at high resolution with ALMA. In the outer regions of \flower, on the other hand, $Q_{\rm eff}$ resembles $Q_{\rm gas}$, with both $\log{Q_{\rm gas}}$ and $\log{Q_{\rm eff}} < -1$. Notably, these are the more gas-rich regions (see \Fig{sigma}).

@@

The sizes of MCCs are only somewhat dependent on the choice of $n_{\rm cut}$, which is a proxy to different tracers used in observations. 
For instance, ......
after excluding the 
most massive molecular structure identified (which correspond to the main disk of \flower). 
The $\sim$150\,pc-scale structure found with a low density threshold correspond to the main disk of \flower 
and breaks down into smaller MCCs at higher density thresholds, with a typical size of $R\simeq\,50$\,pc.
%
The typical mass of MCCs is $10^{6.5}$\,\Msun. Similarly massive molecular structures have been observed in nearby star-forming and starburst galaxies \citep[e.g.,][]{Keto05a, DonovanMeyer13a, Colombo14a, Leroy15a}, and reported in idealized (no stellar feedback) isolated (no cosmological initial conditions) galaxy simulations done at higher resolution (e.g., $l_{\rm cell}\simeq3$\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}). That said, the similar mass range found in MCCs of \flower\ is reassuring --- our results are not far off despite the resolution limit ($l_{\rm cell}\simeq$\,30\,pc). On the other hand, the gas velocity dispersion 
of $\sigma_{\rm gas}\simeq$\,20-10\,\kms is rather insensitive to the adopted density threshold.

>>>>

The most massive structures ($M_{\rm gas}>$10$^7$\,\Msun) correspond
to the molecular disk of the galaxy (see e.g., \Fig{MCC}), which is
identified as a single component at low $n_{\rm cut}$. This main disk
component
%mm2 occupy the top right   [left not right!]
          occupies the top left
corner of \Fig{stellarRatio16}, with a low gas mass fraction of
$f_{\rm gas}< 5\%$. The velocity dispersion of such a structure is
dominated by bulk
%mm2 motions,
rotational motion,
as shown in \Fig{vv}, where we compare the velocity dispersions of MCCs resulting from bulk ($\sigma_{\rm bulk}$) versus non-thermal turbulent motions ($\sigma_{\rm NT}$) in the accreting and quiescent phases of \flower. The accreting phase of \flower displays a disturbed morphology, whereas the quiescent phase displays a disk-like morphology (\Fig{MCC}).
%
Excluding the structure corresponding to the molecular disk, the
velocity dispersions of most MCCs in the accreting phase of \flower
are dominated by turbulent motions, 
%mm2 . On the other hand, [the first hand is missing]
  while  in the quiescent disk-like phase, $\sigma_{\rm bulk}$ is at
  least comparable to $\sigma_{\rm NT}$ for almost half the MCCs.
%mm2 in the quiescent disk-like phase.
%
For these reasons, the most massive MCCs are excluded in the discussion of the dynamics (\Sec{diss}), as they are dominated by large-scale shear; however, we include them in the scaling relation plots (\Fig{larsons_single} and \Fig{alpha16-28}), as would be done for an observational study.

@@

While the virial parameter defined as Equation~\ref{eqn:alpha} is typically used in observational studies, as they are designed to probe structures that are composed of mainly molecular gas only, we calculate the virial parameter of MCCs including the influence from the stellar component (i.e., $\alpha_{\rm vir, tot}$ via Equation~\ref{eqn:alpha_tot}) given the high stellar-to-gas mass ratios of some molecular structures identified (see \Fig{stellarRatio16}).
% 
The biggest MCC identified at low $n_{\rm cut}$ and with the highest stellar-to-gas mass ratio ($\gtrsim$\,1; \Fig{stellarRatio16}) corresponds to the central main 
disk of \flower (see top panels of \Fig{MCC} and \Sec{distribution}). The large amount of stellar mass already assembled in \flower
explains the low {\em total} virial parameter seen in the middle left panel of \Fig{larsons_single}, where the stellar gravitational potential 
influences the overall stability. On the other hand, the MCCs in this phase are mostly stable, with 
virial parameter comparable to some of the molecular structures
observed in the Milky Way,
%mm2 but the former is
     though the MCCs are
     more massive.

@@

%mm2 [fiddled with tenses and plurals]
If constant column density (Larson's third relation) were truly a
fundamental property of molecular clouds,
% or the roughly linear relationship between $\rho$ and $R$ of MCCs.
%mm2 (i.e., 
   and if virial equilibrium were a universal property of molecular clouds with negligible surface pressure, 
one would expect a single point in the $\sigma_{\rm
  gas}^2/R$--$\Sigma$ relation at
% mm a given gas
    that
mass surface density $\Sigma$. 
The variation in $\sigma_{\rm gas}^2/R$ with $\Sigma$ seen here suggests that column density is not a fundamental property of 
molecular clouds
%mm2 [maybe this is excessive, but if we are discussing the issue it
% seems worth citing.]
     in agreement with \citet{Heyer09a}.
Previous studies suggest that the observed mass-size relation of molecular clouds 
is solely a result of $\rho\propto M R^{-3}$ and may be an artifact of the limited range of 
column densities a specific molecular line tracer is sensitive to \citep[see e.g.,][]{Ballesteros02a, Ballesteros11a}.
%For instance, \citet{Ballesteros02a} show that the density
%of clouds ($\rho$) is independent of their size ($R$) when one properly extract these quantities in three-dimensional space instead of
%two-dimensional space; the latter is done in \obs.
%When they project the clouds into two-dimensional space and measure the densities and sizes of clouds for a given
%line tracer, they find $\rho \propto R^{-1}$, which is unsurprising since $N_H$ is just a constant.
%They therefore argue that the mass-size relation is simply a result of $\rho\propto M R^{-3}$
%observed in 2D-space at a fixed column density (cf. an actual physical relation of $M \propto R^{4/3}$ claimed in
%observational studies). % where $\rho$ would be proportional to $R^{-1/2}$
%% since $M$ would then scale as $R^{3}$ (ånd naturally display a linear relationship in $\log M - \log R$ plot).

@@

We investigate possible variations in the dynamics of the molecular structures of \flower 
by adopting different
%mm2 sets
    values
of $n_{\rm cut}$ % in identifying the MCCs
to test the robustness of our results against the choice of density threshold. 
That is, how sensitively are the structure properties, and thus, the 
results presented in \Sec{singless}, dependent on the choice of density thresholds?

@@

In this section, we compare the observable quantities of MCCs obtained from the simulations (e.g., $\sigma_{\rm gas}$, $R$, $M_{\rm gas}$, $\Sigma_{\rm gas}$) with those of molecular gas structures from \obs at lower redshift.
%%
%We exclude the biggest MCC identified in each evolutionary stage, which correspond to the main disk of \flower (i.e., the top right cluster of points in the linewidth-size relation of \Fig{larsons_single} and the top left panel of \Fig{alpha16-28}).
The velocity dispersion of MCCs in \flower is similar to those observed in $z$\ssim2 spatially resolved studies of gas-rich star-forming galaxies, 
spanning a range of $\sigma_{\rm gas}\simeq$\,20$-$80\,\kms (\Fig{larsons_single}; see e.g., \citealt{Swinbank11a}). 
However, the sizes of MCCs are almost a factor of two smaller. This
%mm2 likely suggest
    suggests 
that the sizes of \highz molecular gas structures reported in the literature
are
%mm2
   probably
limited by the spatial resolution. 
%
The velocity dispersions of some MCCs are also comparable to the most turbulent gas clouds observed in the inner Milky way and nearby gas-rich galaxies (e.g., M64; \citealt{Oka01a, Rosolowsky05a, Heyer09a, Leroy15a}), which lie along the locus of $\sigma_{\rm gas}\propto R^{0.56}$. Such high velocity dispersion 
in \flower is dominated by non-thermal energy (\Fig{vv}) and results from the injection of kinetic energy from recent \SF as \flower is assembling its stellar mass (given the high Mach numbers of MCCs; \Sec{distribution}). In fact, by the accreting stage (see \Fig{SFH}a) at $z\simeq$\,7.2, \flower has assembled a stellar mass of $M_\star$\eq7.5\E{9}\,\Msun.
%mm2
   This contrasts with \citet{Krumholz18}, who show that numerical models and analytic arguments point to
   feedback generally not being able to produce velocity dispersions of more than 10--20 km~s$^{-1}$, with
   higher velocities likely being produced by Toomre instability driven radial inflow.

   The pressure of MCCs identified at the highest $n_{\rm cut}$ ranges from $P/k_B\simeq10^{7-9}$\,{\rm K}\,{\rm cm}$^{-3}$, with a median of $10^{7.6}$\,\,{\rm K}\,{\rm cm}$^{-3}$. Such high pressures are comparable to those observed in local ULIRGs; however, the molecular clouds in those are concentrated within their central regions \citep{Downes98a, Sakamoto08a}. In our simulated galaxy, the high pressure MCCs
%mm2 of \flower
are found throughout the disk. This difference likely stems from the different physical mechanisms giving rise to the formation (and thus the nature) of these molecular structures.
%
In local ULIRGs, MCCs are likely formed by shock compression and cloud-cloud collisions that funnel large amounts of gas from the progenitor galaxies towards the central region \citep{Tan00a, Wu18a}. On the other hand, the highly turbulent structures of \flower are the result of extra-planar flows \citep{Kohandel19a} and higher velocity accretion/SN-driven outflows \citep{Gallerani18a}. The presence of extra-planar flows may also be the dominant mode for forming the highly supersonic massive MCCs observed in gas-rich star-forming galaxies at $z$\ssim2. For instance, \citet{Swinbank11a} report ISM pressure of $P/k_B\sim$10$^7$\,K\,\cc and molecular gas mass of $M_{\rm gas}\simeq10^{8-9}$\,\Msun in the star-forming regions of a gas-rich starburst galaxy at \z\eq2.3 based on spatially resolved CO line \obs.

% different evolutionary stages
%mm2 [rearranged]
We do not find any major quantitative differences  (see \Fig{larsons_single}) in the MCCs of \flower with respect to those observed in the nearby Universe between its accreting and starburst phase, which are separated by $\simeq$300\,Myr. However, MCCs in the \SB phase have lower $\alpha_{\rm vir}$ compared to the accreting phase. This is expected for \SF to proceed.

@@

Close resemblance of the $Q_{\rm eff}$ and $Q_{\star}$ maps indicates that contributions from the stellar component play an important role in governing the stability of the MCCs against perturbations. This can be understood since stars in \flower dominate the central part of the galaxy in mass, so their gravitational potential provides a non-negligible contribution to the instability. In fact, this is also illustrated by the differences seen between $\alpha_{\rm vir}$ and $\alpha_{\rm vir, tot}$ 
of the most massive structure (i.e., the gas disk of \flower; see \Fig{larsons_single}).
Similarly, contribution from the thickness of the disk is important in \flower since its disk is warped and has a scale height-to-radius ratio of $h/r_{\rm gal}$\ssim150\,pc/1\,kpc\,$\simeq$\,0.15.
% 
That is, some MCCs are found in regions of $Q_{\rm gas}$\ssim1, which is consistent with the expectation that they correspond to regions of high surface densities that are gravitationally unstable. Note, however, that when including the stabilizing effects due to the stellar potential of \flower and the thickness of its disk (via $\sigma_z$, i.e., $Q_{\rm eff}$; see Equation~\ref{eqn:q_eff}), some of these MCCs are consistent with $Q_{\rm eff} > 1$. 
%
This demonstrates the importance of accounting for stellar contribution and disk thickness when examining the stability of molecular gas structures. This consideration is especially relevant for the relatively evolved and enriched systems at high redshift that are preferentially being imaged at high resolution with ALMA. In the outer regions of \flower, on the other hand, $Q_{\rm eff}$ resembles $Q_{\rm gas}$, with both $\log{Q_{\rm gas}}$ and $\log{Q_{\rm eff}} < -1$. Notably, these are the more gas-rich regions (see \Fig{sigma}).

@@

The sizes of MCCs are only somewhat dependent on the choice of $n_{\rm cut}$, which is a proxy to different tracers used in observations.
%mm2 [needs cleanup]
For instance, ......
after excluding the 
most massive molecular structure identified (which corresponds to the main disk of \flower). 
The $\sim$150\,pc-scale structure found with a low density threshold correspond to the main disk of \flower 
and breaks down into smaller MCCs at higher density thresholds, with a typical size of $R\simeq\,50$\,pc.
%
The typical mass of MCCs is $10^{6.5}$\,\Msun. Similarly massive molecular structures have been observed in nearby star-forming and starburst galaxies \citep[e.g.,][]{Keto05a, DonovanMeyer13a, Colombo14a, Leroy15a}, and reported in idealized (no stellar feedback) isolated (no cosmological initial conditions) galaxy simulations done at higher resolution (e.g., $l_{\rm cell}\simeq3$\,pc in the 48\,kpc box studied by \citealt{Behrendt16a}). That said, the similar mass range found in MCCs of \flower\ is reassuring --- our results are not far off despite the resolution limit ($l_{\rm cell}\simeq$\,30\,pc). On the other hand, the gas velocity dispersion 
of $\sigma_{\rm gas}\simeq$\,20-10\,\kms is rather insensitive to the adopted density threshold.
